The Ranked Nearest Neighbour Interchange space

### 2019-05-06 paper.tex
corrections on why split thm counterexample doesn't work in RNNI

%This part is mainly for myself, not sure whether it is important to have it in the paper
The main reason that the counterexample presented in~\ref{Li1996} works in $\nni$ is following:
Let $T_1$ and $T_2$ be two caterpillar trees.
If each taxon $a$ is repaced by a cherry $(a_1,a_2)$ in both trees $T_1$ and $T_2$, the resulting trees $T_1'$ and $T_2'$ have distance $d_{\nni}(T_1',T_2') = d_{\nni}(T_1,T_2)$ in the $\nni$ graph.
However, this does not work in $\rnni$.
Introducing new cherries means introducing new internal nodes with new ranks.
And a shortest path from $T_1$ to $T_2$ does not give a shortest path from $T_1'$ to $T_2'$, there will be some additional rank swaps needed for the internal nodes of the cherries.
% How do we assign the ranks to the new cherry nodes?
% We could give a precise example here, on 3 or 4 taxa (4 is the max, we can compute RNNI space for up to 8 taxa), because we know how the caterpillar trees with max distance look like.

>>>>

%This part is mainly for myself, not sure whether it is important to have it in the paper
The counterexample presented in~\cite{Li1996} works in $\nni$ is based on following observation:
Let $T_1$ and $T_2$ be two caterpillar trees.
If each taxon $a$ is replaced by a cherry $(a_1,a_2)$ in both trees $T_1$ and $T_2$, the resulting trees $T_1'$ and $T_2'$ have distance $d_{\nni}(T_1',T_2') = d_{\nni}(T_1,T_2)$ in the $\nni$ graph.
When proving~\ref{thm:split_nni},~\cite{Li1996} give an example of two trees $T$ and $R$ on $2n$ taxa that consist of two caterpillar trees each that share the root as parent.
The split given by the edges incident to the root is $1, \ldots, n | n+1, \ldots, 2n$ in both trees $T$ and $R$.
For the given example there is a shortest path in $\nni$ where at first $T$ is transformed to a tree $T'$ with $n$ cherries where each cherry contains one taxon of $\{1, \ldots, n\}$ and one taxon of $\{n+1, \ldots, 2n\}$.
Afterwards, $T'$ is tranformed to a tree $R'$ that contains the same cherries as $T'$.
Finally, all cherries of $R'$ are resolved in order to end up in tree $R$.
For proving that this path is shorter than any path preserving the split $1, \ldots, n | n+1, \ldots, 2n$, the fact that $d_{\nni}(T',R') = d_{\nni}(T|_{\{1, \ldots, n\}}, R|_{\{n+1, \ldots, 2n\}})$ is being used.
However, this does not work in $\rnni$.
Introducing new cherries means introducing new internal nodes with new ranks.
Since there are some rank swaps needed for an $\rnni$ path similar to the shortest $\nni$ path presented above, the path in $\rnni$ that breaks the split $1, \ldots, n | n+1, \ldots, 2n$ is no shortest path.
% How do we assign the ranks to the new cherry nodes?
% We could either create a path where $T'$ and $R'$ have distance equal to $d_{\rnni}(T|_{\{1, \ldots, n\}}, R|_{\{n+1, \ldots, 2n\}})$ and have a lot of additional rank swaps before and after that part OR
% Build $T'$ and $R'$ that contain all the cherries within the least number of moves possible (3n-3) and having a longer path between $T'$ and $R'$.
% Also anything in between these two options is possible. Are we able to disprove every option??
